Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-15 paper/abstract.tex
Merger from overleaf.

Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they either recover or safely shut down as requirements dictate. The DARPA Cyber Assured Systems Engineering program is developing design, analysis, and verification tools that enable systems engineers to design-in cyber-resiliency in a Model-Based Systems Engineering environment. This paper describes automated model transformations that introduce  high-assurance cyber-resiliency components into a system, in particular filters and monitors that prevent malicious input and detect supply chain attacks, respectively. A formal specification defines each component, and is used to verify that the component addresses system level cyber requirements. Implementations are directly synthesized from their specifications, and are automatically proven to preserve the exact meaning of the specifications all the way down to the binary code level. The model transformations are integrated into the Open Source AADL Tool Environment (OSATE). The paper further reports on a case study applying security-enhancing model transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning. In the case study, the model transformations add filters to guard against malformed input, as well as monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

>>>>

Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they either gracefully recover or safely shut down as requirements dictate. The DARPA Cyber Assured Systems Engineering program is developing design, analysis, and verification tools that enable systems engineers to design-in cyber-resiliency in a Model-Based Systems Engineering environment. This paper describes automated model transformations that introduce  high-assurance cyber-resiliency components into a system, in particular filters and monitors that prevent malicious input and detect supply chain attacks, respectively. A formal specification defines each component, and is used to verify that the component addresses system level cyber requirements. Implementations are directly synthesized from their specifications, and are automatically proven to preserve the exact meaning of the specifications all the way down to the binary code level. The model transformations are integrated into the Open Source AADL Tool Environment (OSATE). The paper further reports on a case study applying security-enhancing model transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning. In the case study, the model transformations add filters to guard against malformed input, as well as monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

### 2021-05-15 paper/case-study.tex
Merger from overleaf.

In this section, we apply the BriefCASE tool to the development of a UAV surveillance system in which a UAV receives commands from a ground station to conduct surveillance along a geographical feature such as a river. The on-board mission computer then generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission. The UAV is also given a set of \textit{keep-in} and \textit{keep-out} zones that may constrain its flight path.

We have modeled the system architecture of the UAV in AADL.  It includes a Mission Computer for communicating with the ground station and generating flight plans, and a Flight Control Computer for UAV navigation.  The Mission Computer architecture model includes hardware components such as a processor, memory, and communication devices, as well as software.
%
The initial software architecture model (shown in Figure~\ref{fig:sw-initial}) contains drivers for communication with the Ground Station and Flight Control Computer, a Waypoint Manager component that provides flight plan coordinates to the Flight Control Computer, and the Flight Planner.  The Flight Planner is the open-source UxAS software developed by AFRL~\cite{uxas}.

For this application, UxAS accepts three types of messages.  The \textit{Operating Region} message defines where the UAV can and cannot fly.  The \textit{Line Search Task} message contains a series of waypoints that the UAV should traverse.  The waypoints typically lie along some geographical feature of interest, such as a river or railway.  Note that the UAV may not be able to directly traverse the Line Search Task waypoints due to no-fly zone constraints specified in the Operating Region message.  Anytime after receiving the Operating Region and Line Search Task messages, a ground station can transmit an \textit{Automation Request} message, which instructs UxAS to generate a flight plan that satisfies these constraints.  UxAS passes the flight plan in an \textit{Automation Response} message to the Waypoint Manager.  Because the Flight Control Computer can only process a small number of waypoints at a time, the Waypoint Manager parcels a small number of waypoints corresponding to the current UAV position, and sends them to the Flight Control Computer over a serial connection via the UART Driver.

@@

Within the software model, we have formalized some of the high-level requirements as assume-guarantee contracts.  We perform a formal analysis using the AGREE tool (which is integrated with the BriefCASE environment) to verify that the model satisfies the contracts.  For the initial version of our design, the verification passes.
%
Although we are satisfied with the results of the formal verification using AGREE, we have not yet analyzed the design for cyber-vulnerabilities.  
In BriefCASE, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.  The tools generate a list of new requirements corresponding to cyber vulnerabilities found in the design.  To satisfy these requirements we need to mitigate the vulnerabilities discovered by the analysis by modifying the design.
%
For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in Figure~\ref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring unverified or malicious code (that could potentially be included in the component) will not impact other processes.

@@

The AGREE filter policies for the four UxAS connections are similar, and check that record values contained in the messages are within appropriate ranges.  For example, the Automation Response message filter, which drops message containing malformed flight plans, is defined as shown in Figure~\ref{fig:automation-response-filter}.  Although Latitude, Longitude and Altitude are defined as 64-bit \textit{reals}, the filter ensures only messages containing waypoint values between [-90,9], [-180,180], and [0,15000], respectively.

@@

Similar to the filters, the monitor policies are specified in AGREE. For example, 
%The Response Monitor specification is shown in Figure~\ref{fig:response-monitor}, and 
the Geofence Monitor specification is shown in Figure\ref{fig:geofence-monitor}.

@@

The requirements that have been addressed so far mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on board} the UAV.  But we also want to protect against a compromised Ground Station that could potentially transmit well-formed, but malicious commands. The final cyber-requirement is mitigated by the Attestation transform~\cite{attestation-copland}, which adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.  The Attestation Manager is implemented in CakeML and automatically inserted into the application code base by BriefCASE.  Because the Attestation Gate acts as a filter, the transform automatically generates its complete AGREE specification.%, as shown in Figure~\ref{fig:attestation}.

@@

After transforming the model to address the cyber requirements, the software architecture now appears as shown in Figure~\ref{fig:hardened-sw}.  The components in green were added to the model by way of an automated BriefCASE transform and are critical for mitigating cyber attacks. 
We formally verify the model with AGREE to show that all of the component contracts are satisfied, including the new contracts introduced during the model transformations.
Because it is imperative that these high-assurance component implementations are correct,
we run the SPLAT tool to produce provably-correct code.  
The synthesized code is output to a directory in the build file system with the location specified for each component in the model.  The corresponding correctness proof is used in our assurance case as additional evidence that the vulnerability has been properly mitigated.

@@

The UAV and Ground Station software were deployed on the ODROID XU4 hardware and communicated with each other over Ethernet.  The AMASE flight simulator, representing the Flight Contol Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.  The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the operating region constraints.  Some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior. In addition, we modified the Ground Station to simulate a breach for the attestation test.

We performed an initial set of tests on the un-hardened system (Figure~\ref{fig:sw-initial}) to verify the effectiveness of the malicious code.  The following scenarios were exercised:

\paragraph{Infected Ground Station} In this scenario, one of the UxAS files is modified on the Ground Station. When Ground Station files are modified in such an unauthorized manner, messages sent to the UAV will be rejected by the Attestation Manager.

\paragraph{Malformed Line Search Task message} In this scenario, the Line Search Task message contains a waypoint with a longitude value outside the permitted range.  Such malformed messages could exploit vulnerabilities in the onboard software.  This vulnerability is mitigated by inserting a well-formedness filter.  The filter prevents the Line Search Task message from reaching UxAS.

\paragraph{UxAS vulnerability exploit} In this scenario, Line Search Tasks with greater than 90 waypoints will trigger a vulnerability that will crash UxAS.   When this occurs, UxAS be prevented from generating an Automation Response.  This vulnerability is mitigated by inserting a Response Monitor, which checks to see that UxAS outputs an Automation Response message shortly after receiving an Automation Request.  For this scenario, we chose for the monitor to output a status message, which would then be received by the Ground Station and an appropriate action taken.

\paragraph{UxAS trojan modifies flight plan} In this scenario, a trojan embedded in UxAS will attempt to cause the UAV to fly into the specified keep-out zone by modifying the mission command waypoints in the Automation Response.  However, the UAV will detect that it is being instructed to fly into a keep-out zone and instead return to Home Base.

We repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful, as indicated by the status messages sent from the high-assurance components, shown in Figure~\ref{fig:mitigation-output}.

>>>>

In this section, we apply the BriefCASE tool to the development of a UAV surveillance system in which a UAV receives commands from a ground station to conduct surveillance along a geographical feature. The on-board mission computer then generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission. The UAV is also given a set of \textit{keep-in} and \textit{keep-out} zones that may constrain its flight path.

We have modeled the system architecture of the UAV in AADL.  It includes a Mission Computer for communicating with the ground station and generating flight plans, and a Flight Control Computer for UAV navigation.  The Mission Computer architecture model includes hardware components such as a processor, memory, and communication devices, as well as software.
%
The initial software architecture model (shown in \figref{fig:sw-initial}) contains drivers for communication with the Ground Station and Flight Control Computer, a Waypoint Manager component that provides flight plan coordinates to the Flight Control Computer, and the Flight Planner.  The Flight Planner is the open-source UxAS software developed by AFRL~\cite{uxas}.

For this application, UxAS accepts three types of messages.  The \textit{Operating Region} message defines where the UAV can and cannot fly.  The \textit{Line Search Task} message contains a series of waypoints that the UAV should traverse.  The waypoints typically lie along some geographical feature of interest, such as a river or railway.  Note that the UAV may not be able to directly traverse the Line Search Task waypoints due to no-fly zone constraints specified in the Operating Region message.  Anytime after receiving the Operating Region and Line Search Task messages, a Ground Station can transmit an \textit{Automation Request} message, which instructs UxAS to generate a flight plan that satisfies these constraints.  UxAS passes the flight plan in an \textit{Automation Response} message to the Waypoint Manager.  Because the Flight Control Computer can only process a small number of waypoints at a time, the Waypoint Manager parcels a small number of waypoints corresponding to the current UAV position, and sends them to the Flight Control Computer over a serial connection via the UART Driver.

@@

Within the software model, we have formalized some of the high-level requirements as assume-guarantee contracts.  We perform a formal analysis using the AGREE tool (which is integrated with the BriefCASE environment) to verify that the model satisfies the contracts.  For the initial version of our design, the verification passes.
%
Although we are satisfied with the results of the formal verification using AGREE, we have not yet analyzed the design for cyber-vulnerabilities.  
In BriefCASE, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.  The tools generate a list of new requirements corresponding to cyber vulnerabilities found in the design.  To satisfy these requirements we need to mitigate the vulnerabilities discovered by the analysis by modifying the design.
%
For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring unverified or malicious code (that could potentially be embedded in the component) will not impact other processes.

@@

The AGREE filter policies for the four UxAS connections are similar, and check that record values contained in the messages are within appropriate ranges.  For example, the Automation Response message filter, which drops messages containing malformed flight plans, is defined as shown in \figref{fig:automation-response-filter}.  Although \texttt{Latitude}, \texttt{Longitude} and \texttt{Altitude} are defined as 64-bit \textit{reals}, the filter only passes messages containing waypoint values between [-90,90], [-180,180], and [0,15000], respectively.

@@

Similar to the filters, the monitor policies are specified in AGREE. For example, 
%The Response Monitor specification is shown in \figref{fig:response-monitor}, and 
the Geofence Monitor specification is shown in \figref{fig:geofence-monitor}.

@@

The requirements that have been addressed so far mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on-board} the UAV.  But we also want to protect against a compromised Ground Station that could potentially transmit well-formed, but malicious commands. The final cyber-requirement is mitigated by the Attestation transform~\cite{attestation-copland}, which adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.  The Attestation Manager is implemented in CakeML and automatically inserted into the application code base by BriefCASE.  Because the Attestation Gate acts as a filter, the transform automatically generates its complete AGREE specification.%, as shown in \figref{fig:attestation}.

@@

After transforming the model to address the cyber requirements, the software architecture now appears as shown in \figref{fig:hardened-sw}.  The components in green were added to the model by way of an automated BriefCASE transform and are critical for mitigating cyber attacks. 
We formally verify the model with AGREE to show that all of the component contracts are satisfied, including the new contracts introduced during the model transformations.
Because it is imperative that these high-assurance component implementations are correct,
we run the SPLAT tool to produce provably-correct code.  
The synthesized code is output to a directory in the build file system with the location specified for each component in the model.  The corresponding correctness proof is used in our assurance case as additional evidence that the vulnerability has been properly mitigated.

@@

The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicate with each other over Ethernet.  The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.  The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the operating region constraints.  Some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior. In addition, we modified the Ground Station to simulate a breach for our evaluation of the attestation transform.

We first performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigations in order to verify the effectiveness of the malicious code.  
We repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful.
The following scenarios were exercised (status messages from the high-assurance components in the hardened system are shown in \figref{fig:mitigation-output}):

\paragraph{Infected Ground Station} In this scenario, one of the application files is modified on the Ground Station, which sends the UAV on a trajectory that violates mission requirements. 
On the hardened system, messages sent to the UAV were rejected by the Attestation Manager.

\paragraph{Malformed Line Search Task message} In this scenario, the Line Search Task message contains a waypoint with a longitude value outside the permitted range.  Such malformed messages could exploit vulnerabilities in the on-board software.  
This vulnerability is mitigated by inserting a well-formedness filter.  The filter prevented the Line Search Task message from reaching UxAS.

\paragraph{UxAS vulnerability exploit} In this scenario, Line Search Tasks with greater than 90 waypoints were transmitted from the Ground Station, which triggered a vulnerability that crashes UxAS.   When this occurs, UxAS is prevented from generating an Automation Response.  
This vulnerability is mitigated by inserting a Response Monitor, which checks to see that UxAS outputs an Automation Response message shortly after receiving an Automation Request.  For this scenario, we chose for the monitor to output a status message, which would then be received by the Ground Station and an appropriate action taken.

\paragraph{UxAS trojan modifies flight plan} In this scenario, a trojan embedded in UxAS will attempt to cause the UAV to fly into the specified keep-out zone by modifying the mission command waypoints in the Automation Response.  
On the hardened system the Geofence Monitor detected that it was being instructed to fly into a keep-out zone and instead returned the UAV to Home Base.

### 2021-05-15 paper/conclusion.tex
Merger from overleaf.

The DARPA CASE program is creating an MBSE environment for designers to integrate cyber-vulnerability analysis and mitigation. The resulting BriefCASE tool suite includes analyses to add cyber requirements to systems and architectural transformations on systems to satisfy cyber requirements. The filter transformation prevents malformed data from being propagated downstream while the monitor transformation checks temporal properties to detect malicious behavior. The components are specified by corresponding auto-generated AGREE contracts that only require policy annotations taken from the cyber requirements. The specifications are automatically synthesized to CakeML and that synthesis preserves the meaning of the specifications.

The BriefCASE analyses and transformations are applied to a full-scale case study on the Air Force Research Laboratory's OpenUxAS services requiring several filters and monitors to meet cyber-requirements. 
%A video shows the synthesized system running on the ODROID platform with seL4. 
The size and scale of the study suggests BriefCASE meets the complexity demands of real-world design.

Ongoing work is applying BriefCASE in the Collins Common Avionics Architecture System (CAAS)~\cite{caas}, to protect against \emph{Automatic Dependent Surveillance-Broadcast} spoofing. That work adds support for uninterpreted functions. Other ongoing work is mechanizing the synthesis proof in HOL4 and lifting the proof results to infinite streams.

>>>>

The DARPA CASE program is creating tools for systems engineers to integrate cyber-vulnerability analysis and mitigation. The resulting BriefCASE tool suite includes analysis tools for generating cyber requirements, cyber resiliency tools for addressing the requirements, verification tools for ensuring design correctness, and synthesis tools for generating provably correct code. Several of the BriefCASE transforms (filter, monitor, gate) insert components into the model whose behavior can be formally specified in the AGREE language.
The SPLAT tool can then automatically generate CakeML implementations for these components, along with proofs of correctness for assurance that the implementation satisfies the specification.

BriefCASE was applied to a full-scale case study using the Air Force Research Laboratory's OpenUxAS software, exercising a range of built-in cyber resiliency mitigations to meet cyber-requirements. 
The size and scale of the study suggests BriefCASE meets the complexity demands of real-world design.

We are currently in the process of applying BriefCASE to the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47.  Other ongoing work includes adding support for uninterpreted functions, mechanizing the synthesis proof in HOL4 and lifting the proof results to infinite streams.

### 2021-05-15 paper/intro.tex
Merger from overleaf.

The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
%\footnote{The
%  views expressed are those of the authors and do not reflect the
%  official policy or position of the Defense Advanced
%  Research Projects Agency (DARPA) or the U.S. Government.}
The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.

@@

Another novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, a verified compiler implementation for the functional programming language ML \cite{cakeml}. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

The approach is motivated, and illustrated, in a simple example in Section~\ref{sec:example}. A case study applying these transformations to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented in Section~\ref{sec:case-study}. Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS. The case study system is significantly more complex than the example and shows the viability of the modeling approach to full-scale industrial design.

The BriefCASE tools are open source and publicly available \cite{fmide}, as is the motivating example \cite{repo}, and the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.  Videos demonstrating the use of the BriefCASE tools to build the UAV example presented in Section~\ref{sec:case-study} are available at \cite{case}.

>>>>

The DARPA Cyber Assured Systems Engineering (CASE) program is targeted at developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
%\footnote{The
%  views expressed are those of the authors and do not reflect the
%  official policy or position of the Defense Advanced
%  Research Projects Agency (DARPA) or the U.S. Government.}
We have developed a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

@@

Another novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, a verified compiler implementation for the functional programming language ML \cite{cakeml}. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR-generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

The approach is motivated, and illustrated, in a simple example in Section~\ref{sec:example}. Contract specification in the AGREE language is presented in Section~\ref{sec:agree}, followed by a description of the synthesis pathway in Section~\ref{sec:synthesis}. A case study applying these transformations to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented in Section~\ref{sec:case-study}. Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS. The case study system is significantly more complex than the simple example and shows the viability of the modeling approach to a full-scale industrial design.

The BriefCASE tools are open source and publicly available \cite{fmide}, as is the motivating example \cite{repo}, and the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.  Videos demonstrating the use of the BriefCASE tools to build the UAV example presented in Section~\ref{sec:case-study} are also available \cite{case}.
